---
title: "ds202_lab3_team"
author: "HanSun, XinyiZhu"
date: "2020/2/19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggmosaic)
library(readxl)
library(dplyr)
library(tidyr)
```
```{r}
dat <- readxl::read_xls('GSS.xls')
names(dat) <- c('year','id','marital','sex','party','happiness','ballot')
dat <- dat %>% drop_na()
str(dat)
```
1. How does the happiness of a respondent relate to the marriage status?
```{r}

dat <- dat %>% mutate(happy = case_when(
  happiness == "Not too happy" ~ "Not happy",
  happiness == "Pretty happy"  ~ "Happy",
  happiness == "Very happy"  ~ "Happy",
  happiness == "Don't know" ~ "No answer",
  happiness == "No answer" ~ "No answer",
  happiness == "Not applicable" ~ "No answer"))
dat %>% 
  filter(!is.na(marital)) %>%
  ggplot(aes(x= marital, fill = happy)) + geom_bar(position = 'fill')
```

For married people, about 80% said they were happy. About 70% of the separated people said they were happy. It is, therefore, easier for married people to be happy. About 25% of the separated people said they were not happy. And less than 10% of married people say they are not happy.

2. Does the sex of the respondent affect the relationship you found in Q1?
```{r,fig.width=12, fig.height=2}
dat %>% 
  filter(!is.na(sex)) %>%
  ggplot(aes(x= marital, fill = happy)) + geom_bar(position = 'fill')+facet_grid(~sex)
```

In fact, the gender of the respondents did not affect the findings in Q1.

3. How does the happniess of a respondent relate to the political party affiliation?
```{r}
dat[is.na(dat$`party`), ]
dat1 <- dat[!is.na(dat$'party'), ]

ggplot(dat1, aes(x=party, fill=happiness)) + geom_bar(position='fill') + coord_flip()
```

For people whose party "Don't know" has the largest part of proportion of "very happy"and also has the largest part of "not too happy". For people who feel "very happy"and "not applicable", parties which are not straight democracyand republican and strong democracy and republican are weighted similar. For people who feel "Pretty happy", not straight democracy weighted the largest proportion. For people who feel not too happy, people whose party is not straight dempcracy is weighed higher than other parties.For people who no answer, people whose party is strong republican is weighed less than other parties. 


4. Raise four questions that can be answered by the GSS dataset. Extract appropriate data from the website, explore, and answer the questions.
```{r}
# Question 1 #
## How does the year of a respondent relate to the marriage status?
dat <- dat %>% mutate(time = case_when(
  year < 1980 ~ "1970s",
  year < 1990 ~ "1980s",
  year < 2000 ~ "1990s",
  year < 2010 ~ "2000s",
  year < 2020 ~ "2010s"))
ggplot(dat, aes(x = marital, fill = time)) + geom_bar()
```

People in the 1970s had the lowest probability of divorce. For other marital statuses, the proportion of people in each age is similar.

```{r}
# Question 2#
## How does the sex of a respondent relate to the happiness?
ggplot(dat1, aes(x=happiness, fill=sex)) + geom_bar(position='fill')

```

For each options of happiness, female seem to be more emotional than male since the proportion of female answers are larger than 0.5.


```{r,fig.width=12, fig.height=2}
# Question 3
## How does the ballet of a respondent relate to the public party?
ggplot(dat1, aes(x=party, fill=ballot)) + geom_bar(position='fill')
```

Proportions of ballot a and b are pretty similar in all parties expect "Don't know". Ballot d has the smallest amount proportions in all parties. There are lots of "Not applicable" throughout the dataframe, so we need to consider the reliability of the conclusion.

```{r}
# Question 4
## How does the year of a respondent relate to the political party?
dat1 <- dat1 %>% mutate(time = case_when(
  year < 1980 ~ "1970s",
  year < 1990 ~ "1980s",
  year < 2000 ~ "1990s",
  year < 2010 ~ "2000s",
  year < 2020 ~ "2010s"))
ggplot(dat1, aes(x = happy, fill = time)) + geom_bar()
```

People,who are 20 years old, are not likely to give an answer.  People of 30 and 40 years old are the happiest part of people in this dataframe. People of different ages were equally likely to choose to unhappy.
